---
title: "Spotify ML 2 project"
author: "Weronika Motkowska"
date: "2023-01-16"
output: rmdformats::material
---

# Audio features

The idea for this project has been inspired by paper by Sciandra, Mariangela; Spera, Irene; 2020/01/01; "A Model Based Approach to Spotify Data Analysis: A Beta GLMM"; 10.2139/ssrn.3557124; SSRN Electronic Journal.

The aim of this paper is to investigate if audio features from Spotify can be considered as determinants of the stream popularity of songs. In here the extracted data will consider different artist and the explained variable will be bounded popularity which can have values from 0 to 100 however the values are extracted the way the boundary values are not really possible so the explained variable can be considered as continuous.

```{r warning=FALSE, message=FALSE}
library(rvest)
library(tidyverse)
library(spotifyr)
library(knitr)
library(ggplot2)
```

```{r echo=FALSE}
Sys.setenv(SPOTIFY_CLIENT_ID = "7279343fb60242fe8ee1194a4f7fe5df")
Sys.setenv(SPOTIFY_CLIENT_SECRET = "10b0750b71d34615bf5e3396097b9b23")
access_token <- get_spotify_access_token()
```

```{r}
get_audio_features <- function(artist_name, artist_id, genre){
  
  artist <- search_spotify(artist_name, type = "artist", market = "US", limit = 1)
  top <- get_artist_top_tracks(artist_id, 
                               market = "US") %>% 
    select(name, id, popularity)
  
  top.features <- data.frame()
  for (id in top$id){
    top.features <- get_track_audio_features(id) %>% 
      bind_rows(top.features)
  } 
  
  top <- top %>% 
    left_join(top.features) %>% 
    select(-c(15:18, 20)) %>% 
    mutate(followers.total = artist$followers.total) %>% 
    mutate(genre = genre)
  
  rm(artist, 
     top.features)
  
  return(top)
}
```

```{r}
# artists <- get_genre_artists(genre = "pop", market = "US", limit = 20) %>%
#              mutate(genre = "pop") %>%
#            bind_rows(get_genre_artists(genre = "rap", market = "US", limit = 20) %>% mutate(genre = "rap")) %>%
#            bind_rows(get_genre_artists(genre = "rock", market = "US", limit = 20) %>% mutate(genre = "rock")) %>%
#            select(name, id, genre)
```

```{r message=FALSE, warning=FALSE}
# artist.data <- data.frame()
# for (i in 1:nrow(artists)){
#   artist_name <- artists[i, 1][[1]]
#   artist_id   <- artists[i, 2][[1]]
#   genre <- artists[i, 3][[1]]
#   artist.data <- bind_rows(artist.data, get_audio_features(artist_name, artist_id, genre))
# }
# 
# save(artist.data, file = "data")
```


```{r}
load("data")
str(artist.data)
data <- artist.data %>% select(-c(name, id, genre)) %>% scale() %>% data.frame()
```

# Simple analysis

```{r}
ggplot(artist.data, aes(popularity, fill = genre, colour = genre)) +
  geom_density(alpha = 0.2, na.rm = TRUE) 
```

```{r}
options(scipen = 8)
model.lm <- lm(popularity ~ ., data)
summary(model.lm)
```
The top songs in our catalog are around 86 score in popularity.

Danceability and loudness have significant positive impact on popularity. 
The hypothesis is that the song is more popular if it is catchy and suitable for dancing on TikTok. 

The simple analysis shows that popularity of the artist significantly influences popularity score of the song.
The longer the song the less popular it is, with duration having negative sign next to it. 

What is interesting is that the song does not have to be energetic to be popular and danced to. The positiveness (valence) is also considered not popular.

Other charasteristics, which are more technical, seem not to have much of importance.


